Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2016-03-02 draft.tex
simplify prob in war lotter from p1 to p

For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_i^t \equiv (p_1 (W_i^t-\zeta_i), \left(1-p_1\right)( L_i^t-\zeta_i))$. $\Omega^t \equiv (\omega_g^t, \omega_s^t).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$. Note that the structure of the war lottery along with the assumption that the outside actors influence the payoffs only and not the probabilities of winning imply that any way in which the patron influences the payoffs from war can be counteracted by the international community through its own investments.} Players are assumed to approach war lotteries as expected values.

@@

\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_{1i}) + L_i^0(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}

@@

\item  \textit{$\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}

@@

	\item As long as $R_p^t \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_{1s} + L_s^{t-1}(1-p_{1s})\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_{1s} + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_{1s} + L_s^{t-1}(1-p_{1s})\right)$ in $q_s^t$. Similar to case two, the outcome is that investments in $w_s^t$ and $q_s^t$ are zero.

If $R_p^t > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_{1s} + L_s^{t-1}(1-p_{1s})\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$. Fight becomes the unique best response for the secessionists.
\end{enumerate}
\end{definition}

@@

\begin{lemma}
$Q_i^t \geq L_i^t$ and $\frac {Q_i^t}{1-\delta} \geq  -\zeta_i+\frac{W_i^t(p_{1i}) + L_i^t(1-p_{1i})}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$. 
\end{lemma}

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e. $q_s^{t+1} = Q_s^t - \mu -\sigma$ \emph{and} $p_{1s}\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$). The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
\end{proposition}

@@

\item \hangindent=1cm van Houten, Pieter. 1998. "The role of a minority's reference state in ethnic relations." \emph {European Journal of Sociology / Archives EuropŽennes de Sociologie} 39 (01): 110-146.

@@

{\bf Proof of Lemma 1} \\
Recall that the upper case notation represents payoffs gross of investments by the outside players. Without loss of generality, consider the incentives to deviate for player $s$ in period $t$ given that the other three players play their equilibrium actions. Given the stationary equilibrium actions, playing Status Quo will lead to $Q_s^t \ \forall t$, so the continuation value is $\frac{Q_s^t}{1-\delta}$. Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_s^t$ in each period, is $\frac{L_s^t}{1-\delta}$. The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$. Thus the continuation value is $-\zeta_s +\frac{W_s^t(p_1)+ L_s^t(1-p_1)}{1-\delta}$.

@@

Since the patron is willing to invest up to $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$, when $\frac{\nu}{1 -\de} \geq \frac{\lambda + \beta + \mu}{1 -\de} -\mu$, player $c$'s willingness to counter the patron's investments is sufficient to prevent the \emph{recognition} outcome. A slightly weaker version of this inequality is stated as Restriction (4) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 4} \\
The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ} -R_p^{WAR}$. For player $c$ it is $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de} + R_c^{SQ} -R_c^{WAR}$. As noted above, $R_c^{SQ} =0$. So $c$ is willing to invest up to $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting.

The patron is willing to invest up to $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive. Thus we must only examine the case where this quantity is positive.

We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de} -\mu$ as this is most restrictive. In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu> 0$ for an upper bound on the amount the patron is willing to invest to instigate fighting.

If $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  < \omega_s^t\left(R_p^t\right) - Q_s^t$, the patron's investment in $\omega_s^t$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.

On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g. counter an addition to $\omega_s^t$ with an investment to $q_s^t$---as long as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq \omega_s^t\left(R_p^t\right) - Q_s^t$.

This can be expanded as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_s^{t-1}+R_p^t)p_{1s} + L_s^{t-1}(1-p_{1s}) - Q_s^{t-1} +\mu = \omega_s^{t-1} - Q_s^{t-1} +R_p^t p_{1s} + \mu$.

Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \geq R_p^t p_{1s}$.

Since $R_p^t \leq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$, if we can show $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  - \mu \geq p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} - p_{1s} \mu$, we will have the guarantee that player $c$ will counterinvest after any effort by the patron to instigate fighting, thereby making $p$'s optimal investment in fighting zero.

Multiplying Restriction (3) by $-(1-p_{1s})$, we have $\frac{-\beta \cdot (1-p_{1s})}{1-\de} -\mu (1-p_{1s}) \geq \frac{-\alpha \cdot (1-p_{1s})}{1-\de}$. Using this with Restriction (4) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1-\de} -\mu \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu$. Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu > p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu$. \hfill $\blacksquare$
\\
\\
{\bf Proof of Proposition 2} \\
Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$. As the only alteration is a reduction in $Q_s^t$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed. As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.

@@

Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 - \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1-\de}$. For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{-\nu p_{1s} + \beta (1 - p_{1s})}{1-\de}$. Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
\[
  \textstyle \frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1-\de} + \frac{\de^m}{1-\de} \alpha > \frac{-\nu p_{1s} + \beta (1-p_{1s})}{1-\de} -  \frac{\de^m}{1-\de} \beta
\]

>>>>

For probabilities $p$ of outright victory and $1-p$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_i^t \equiv (p (W_i^t-\zeta_i), \left(1-p\right)( L_i^t-\zeta_i))$. $\Omega^t \equiv (\omega_g^t, \omega_s^t).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$. Note that the structure of the war lottery along with the assumption that the outside actors influence the payoffs only and not the probabilities of winning imply that any way in which the patron influences the payoffs from war can be counteracted by the international community through its own investments.} Players are assumed to approach war lotteries as expected values.

@@

\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_i) + L_i^0(1-p_i)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}

@@

\item  \textit{$\nu\cdot p_s \geq \lambda \cdot p_s + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}

@@

	\item As long as $R_p^t \leq \frac{1}{p_s} \left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_s + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)$ in $q_s^t$. Similar to case two, the outcome is that investments in $w_s^t$ and $q_s^t$ are zero.

If $R_p^t > \frac{1}{p_s}\left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$. Fight becomes the unique best response for the secessionists.
\end{enumerate}
\end{definition}

@@

\begin{lemma}
$Q_i^t \geq L_i^t$ and $\frac {Q_i^t}{1-\delta} \geq  -\zeta_i+\frac{W_i^t(p_i) + L_i^t(1-p_i)}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$. 
\end{lemma}

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e. $q_s^{t+1} = Q_s^t - \mu -\sigma$ \emph{and} $p_s\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$). The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
\end{proposition}

@@

\item \hangindent=1cm van Houten, Pieter. 1998. "The role of a minority's reference state in ethnic relations." \emph {European Journal of Sociology / Archives Europennes de Sociologie} 39 (01): 110-146.

@@

{\bf Proof of Lemma 1} \\
Recall that the upper case notation represents payoffs gross of investments by the outside players. Without loss of generality, consider the incentives to deviate for player $s$ in period $t$ given that the other three players play their equilibrium actions. Given the stationary equilibrium actions, playing Status Quo will lead to $Q_s^t \ \forall t$, so the continuation value is $\frac{Q_s^t}{1-\delta}$. Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_s^t$ in each period, is $\frac{L_s^t}{1-\delta}$. The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p$, and $L_s$ forever with probability $1-p$. Thus the continuation value is $-\zeta_s +\frac{W_s^t(p)+ L_s^t(1-p)}{1-\delta}$.

@@

Since the patron is willing to invest up to $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$, when $\frac{\nu}{1 -\de} \geq \frac{\lambda + \beta + \mu}{1 -\de} -\mu$, player $c$'s willingness to counter the patron's investments is sufficient to prevent the \emph{recognition} outcome. A slightly weaker version of this inequality is stated as Restriction (4) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 4} \\
The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_s - \alpha (1-p_s)}{1 -\de} +R_p^{SQ} -R_p^{WAR}$. For player $c$ it is $\frac{-\nu p_s + \beta (1-p_s)}{1 -\de} + R_c^{SQ} -R_c^{WAR}$. As noted above, $R_c^{SQ} =0$. So $c$ is willing to invest up to $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}$ toward preventing fighting.

The patron is willing to invest up to $\frac{\lambda p_s - \alpha (1-p_s)}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive. Thus we must only examine the case where this quantity is positive.

We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de} -\mu$ as this is most restrictive. In this only case of interest, we have $\frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1 - \de} -\mu> 0$ for an upper bound on the amount the patron is willing to invest to instigate fighting.

If $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  < \omega_s^t\left(R_p^t\right) - Q_s^t$, the patron's investment in $\omega_s^t$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.

On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g. counter an addition to $\omega_s^t$ with an investment to $q_s^t$---as long as $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  \geq \omega_s^t\left(R_p^t\right) - Q_s^t$.

This can be expanded as $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_s^{t-1}+R_p^t)p_s + L_s^{t-1}(1-p_s) - Q_s^{t-1} +\mu = \omega_s^{t-1} - Q_s^{t-1} +R_p^t p_s + \mu$.

Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu \geq R_p^t p_s$.

Since $R_p^t \leq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1 - \de} -\mu$, if we can show $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  - \mu \geq p_s \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1 - \de} - p_s \mu$, we will have the guarantee that player $c$ will counterinvest after any effort by the patron to instigate fighting, thereby making $p$'s optimal investment in fighting zero.

Multiplying Restriction (3) by $-(1-p_s)$, we have $\frac{-\beta \cdot (1-p_s)}{1-\de} -\mu (1-p_s) \geq \frac{-\alpha \cdot (1-p_s)}{1-\de}$. Using this with Restriction (4) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_s - \beta (1-p_s)}{1-\de} -\mu \geq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$. Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu > p_s \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$. \hfill $\blacksquare$
\\
\\
{\bf Proof of Proposition 2} \\
Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$. As the only alteration is a reduction in $Q_s^t$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed. As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.

@@

Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 - \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_s - \alpha (1-p_s)}{1-\de}$. For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{-\nu p_s + \beta (1 - p_s)}{1-\de}$. Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
\[
  \textstyle \frac{\lambda p_s - \alpha (1-p_s)}{1-\de} + \frac{\de^m}{1-\de} \alpha > \frac{-\nu p_s + \beta (1-p_s)}{1-\de} -  \frac{\de^m}{1-\de} \beta
\]
